---
title: "Week 04 Notes"
author: "Haley Poppinga"
date: "2025-09-16"
output: github_document
---

Lecture In-Class: Intro to {dplyr}
 - a set of verbs for manipulating data:
  - filtering rows, selecting columns, grouping by objects, getting summary statistics, mutating columns, removing NAs

Homework 
1. Practice with dplyr

[dplyr cheat sheet.](https://raw.githubusercontent.com/rstudio/cheatsheets/main/data-transformation.pdf) Download this now to use as a resource.


# Review

1. Where do I change the font size of my x and y labels?
Theme

2. How do I save my plot?
ggsave


# Intro to the {dplyr} package (part of the TidyVerse)
Don't need to load dplyr. Just TidyVerse

Illustrations from the Openscapes blog [Tidy Data for reproducibility, efficiency](https://www.openscapes.org/blog/2020/10/12/tidy-data/), and collaboration by Julia Lowndes and Allison Horst


# dplyr: verbs for manipulating data
[verbs](libs/images/dplyr_verbs.PNG)

filtering extracts rows
select() extracts columns
arrange() 
mutate() adds new columns, same length of rows but longer
group_by() %>% summarize() makes shorter


# Penguin data again.

```{r}
### Today we are going to plot penguin data ####
### Created by: Haley Poppinga #############
### Updated on: 2025-09-17 ####################


#### Load Libraries ######
library(palmerpenguins)
library(tidyverse)
library(here)


### Load data ######
# The data is part of the package and is called penguins
glimpse(penguins) 



```


# Filter

dplyr::filter()
call the package name first
filter(df, type == "otter" & site == "blah")

== reads as is exactly equal to


#Filter 
### Extract rows that meet some critera

```{r}
#filter(.data = DATA, ...)
```


* DATA = Data frame to transform  

* ... = One or more criteria  
filter() returns each row for which the criteria is TRUE


# filter only the female penguins
### As always, exact spelling and capitalization matters

Before filtering
```{r}
head(penguins)
```


# filter only the female penguins
### As always, exact spelling and capitalization matters

After filtering
```{r}
filter(.data = penguins, sex == "female" )
```
is sex exactly equal to female?
argument = .data

# filter

```{r}
filter(.data = penguins, 
       sex == "female" ) #filter keeps every TRUE
```

One = sets an argument in the function  

Two == reads as "is exactly equal to." It is a question that returns a TRUE or FALSE.  Here, filter keeps every TRUE


# A list of logical expressions 

[logicals](libs/images/logical_test.PNG)

Test Meaning
x < y less than
x > y greater than
== equal to (if you only have one thing to pull out)
x <= y less than or equal to
x >= y greater than or equal to
x != y not equal to (! can be used with anything)
x %in% y in group membership (pull out more than one thing, e.g is the species within the vector)
is.na(x) is missing
!is.na(x) is not missing (is it not an NA, if so retain it)

Note: these used universally in all coding languages


# How would I use filter to...

1. Penguins measured in the year 2008?
```{r}
filter(.data = penguins, 
       year == 2008)
```


2. Penguins that have a body mass greater than 5000
```{r}
filter(.data = penguins, 
       body_mass_g > 5000)
```




# Think, pair, share.  

Spend 2 minutes to *think* about it. Spend 2 minutes *paired* with a neighbor to discuss your answers. *Share* with the class.


# Common mistakes

## Using = instead of ==

```{r}
#filter(.data = penguins,
#  sex = "females")

filter(.data = penguins, 
       sex == "females")
```


## Forgetting quotes  

```{r}
#filter(.data == penguins, 
    #    sex == females) #for things that are characters

filter(.data = penguins, 
       sex == "females")
```
 



# Filter with multiple conditions

Select females that are also greater than 5000 g
comma means "and" --> do this and that
```{r}
filter(.data = penguins, sex == "female", body_mass_g >5000)

```



[operators](libs/images/operators.PNG)
Boolean operators
a & b --> and
a | b --> or
!a --> not


# Default for filter is **&**

These do the same exact thing
```{r}
filter(.data = penguins, sex == "female", body_mass_g >5000)

filter(.data = penguins, sex == "female" & body_mass_g >5000)

```



# Think, pair, share

Use filter and boolean logical to show:
#always need quotes for calling a character

1. Penguins that were collected in *either* 2008 *or* 2009
```{r}
filter(.data = penguins, year == 2008 | year == 2009)
```


2. Penguins that *are not* from the island Dream
```{r}
filter(.data = penguins, island != "Dream")
```


3. Penguins in the species Adelie and Gentoo
```{r}
filter(.data = penguins, species %in% c("Adelie","Gentoo"))
```



# Common mistakes

### Collapsing multiple tests into one 
penguins between 3000 and 5000 g
```{r}
#Wrong
#filter(.data = penguins,
 #      3000 < body_mass_g <5000) #

#Right
filter(.data = penguins, 
       body_mass_g <5000, #need to recall both 
       body_mass_g >3000) #can only have one logical statement per question

```

### Using multiple tests instead of %in%
penguins in dream and biscoe

```{r}
# Wrong
#filter(.data = penguins,
 #      island == "Dream", #<<
  #     island == "Biscoe") #<<

# Right
filter(.data = penguins,
       island %in% c("Dream","Biscoe")) #pull from within

```



[syntax](libs/images/syntax.PNG)
Common syntax
every dplyr verb function follows 


# mutate: add new columns

# mutate
[mutate](libs/images/mutate.PNG)


# mutate
*Create new columns*
mutate(.data,...)
* DATA = data frame to transform
* ... = columns to make
Add a new column converting body mass in g to kg and save the new dataframe

```{r}
data2<-mutate(.data = penguins, 
       body_mass_kg = body_mass_g/1000) #<<

View(data2)

```



### Change multiple columns at once
```{r}
data2<-mutate(.data = penguins, 
       body_mass_kg = body_mass_g/1000, #new column
       bill_length_depth = bill_length_mm/bill_depth_mm) #new column

# both data frames now within data2
# first line is convert mass to kg 
# second line is calculate the ratio of bill length to depth

View(data2)

```

## Mutating multiple columns at once based on some criteron (i.e. mutate all the numeric columns in one line)?

On your own, look up
* *mutate_if()* asks if its logical, eg log or squareroot everything
* *mutate_at()* selects specific variables
* *mutate_all()* mutates every single column in whatever way you want it to

[Some examples here](https://dplyr.tidyverse.org/reference/mutate_all.html)


[ifelse](libs/images/ifelse.PNG)
ifelse()
Do conditional tests within mutate()

ifelse(TEST,
      VALUE_IF_TRUE,
      VALUE_IF_FALSE)

* TEST = a logical test
* VALUE_IF_TRUE = what happens if test is true
* VALUE_IF_FALSE = what happensif test is false



# mutate with ifelse

```{r}
data2<- mutate(.data = penguins,
               after_2008 = ifelse(year>2008, "After 2008", "Before 2008"))
# if the word is greater than 2008, what 
View(data2)
```



# Think, pair, share

1. Use mutate to create a new column to add flipper length and body mass together
```{r}
data2<-mutate(penguins, 
      length_body_mass = flipper_length_mm + body_mass_g) 
view(data2)
```


2. Use mutate and ifelse to create a new column where body mass greater than 4000 is labeled as big and everything else is small 
```{r}
data2<- mutate(.data = penguins,
               big = ifelse(body_mass_g>4000, "big","small"))
view(data2)
```



# What if you have multiple verbs that you want to do to the same dataframe?



# The pipe  
# %>%

## The "pipe" says "and then do"

Dataframe %>%  # select the dataframe **and then** 
> verb1() %>%  # do verb 1 **and then** 

>verb2()      # do verb 2


# %>%


me %>%
  wake_up(time = "6:00") %>%
  music(turn = "on", to = "Beyonce") %>%
  shower(wash_hair = TRUE) %>%
  get_dressed(pants = TRUE, shirt = TRUE) %>%
  leave_house(car = TRUE)


# What about this new *native* pipe that I keep seeing on stack exchange?

# %>%   (need to load TidyVerse in order for it to run, more universal than native)
# vs  
# |> 

**magrittr** (from the tidyverse) vs **native** (new pipe that comes with base R so you don't need to load any libraries) 

In this class I stick with %>%, but you are welcome to learn and use |>, with the understanding that there will be some minor differences in the code.

[Please Read Here!](https://r4ds.hadley.nz/data-transform#sec-the-pipe)


# %>%

Filter only female penguins and add a new column that calculates the log body mass  
When you use %>% the dataframe carries over so you don't need to write it out anymore
* pipe and then new line, pipe and then new line, everything is connected

```{r}
penguins %>% # use penguin dataframe
  filter(sex == "female") %>% #select females
  mutate(log_mass = log(body_mass_g)) #calculate log biomass

```


# Select

Use select() to select certain columns to remain in the dataframe

```{r}
penguins %>% # use penguin dataframe
  filter(sex == "female") %>% #select females
  mutate(log_mass = log(body_mass_g)) %>% #calculate log biomass
  select(species, island, sex, log_mass)

```


# Select

You can also use select() to rename columns.  
Here, we are renaming species to have a capital S

```{r}
penguins %>% # use penguin dataframe
  filter(sex == "female") %>% #select females
  mutate(log_mass = log(body_mass_g)) %>% #calculate log biomass
  select(Species = species, island, sex, log_mass)

```

# Summarize 
 Computer a table of summarized data  
  Calculate the mean flipper length (and exclude any NAs)
 
```{r}
 penguins %>% # 
  summarise(mean_flipper = mean(flipper_length_mm, na.rm=TRUE)) #<<
```


Calculate mean and mean flipper length
 
```{r}
 penguins %>% # 
  summarise(mean_flipper = mean(flipper_length_mm, na.rm=TRUE), #<<
            min_flipper = min(flipper_length_mm, na.rm=TRUE))   #<<
```


# group_by
You can summarize values by certain groups.  
group_by() by itself doesn't do anything, but it is powerful when put before summarize.
* use to make bar plots and standard errors in summary statistics

Let's calculate the mean and max bill length by island

```{r}
penguins %>%
  group_by(island) %>% #<<
  summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            max_bill_length = max(bill_length_mm, na.rm=TRUE))
```


# group by
Groupe by multiple things
Group by both island **and** sex

```{r}
penguins %>%
  group_by(island, sex) %>% #<<
  summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            max_bill_length = max(bill_length_mm, na.rm=TRUE))
#sumarise() has grouped output by 'island'. you can override using the 
# .groups argument
```



# remove NAs

drop_na() 
  - drops rows with NAs from a specific column

drop all the rows that are missing data on sex
```{r}
penguins %>%
  drop_na(sex) #<<

```

# remove NAs

drop_na() 
  - drops rows with NAs from a specific column
  - only use if you want to drop every single missing data from sex

drop all the rows that are missing data on sex
calculate mean bill length by sex

```{r}
penguins %>%
  drop_na(sex) %>%
  group_by(island, sex) %>% #cleans up NAs
  summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))

```

# Pipe into a ggplot
You can connect your *data wrangling* to a ggplot with the pipe (you won't need to call the dataframe in ggplot if you pipe to it).
* pipe straight into ggplot

Drop NAs from sex, and then plot boxplots of flipper length by sex

```{r, fig.height=4}
penguins %>%
  drop_na(sex) %>%
  ggplot(aes(x = sex, y = flipper_length_mm)) + #go back to + when in ggplot
  geom_boxplot()
#if you assign something to this it'll be a plot not a data set

```


# Before homeworks... "Totally Awesome R package of the day"


library(devtoosls) # load the development tools library
devtools::install_github("jhollist/dadjoke")

library(dadjoke)
dadjoke()



# Homework

Write a script that:

1. calculates the mean and variance of body mass by species, island, and sex without any NAs

1. filters out (i.e. excludes) male penguins, then calculates the log body mass, then selects only the columns for species, island, sex, and log body mass, then use these data to make any plot. Make sure the plot has clean and clear labels and follows best practices. Save the plot in the correct output folder.

Include both part 1 and part 2 in your script and push it to github in the appropriate folders.

```{r}
penguins %>%
  drop_na(species, island, sex, body_mass_g) %>% #removes NAs for species, island, and sex
  group_by(species, island, sex, body_mass_g) %>% #body mass by species, island, and sex
  summarise(mean_body_mass = mean(body_mass_g, na.rm = TRUE), #remove NAs from body mass
            variance_body_mass = var(body_mass_g, na.rm = TRUE)) %>% #calculate mean and variance
  filter(sex != "male") %>% #filter out males
  mutate(log_mass = log(body_mass_g)) %>% #calculate log body mass
  select(species, island, sex, log_mass) %>% #selects only the columns for species, island, sex, and log body mass
```

#Plot
  
```{r}
ggplot(penguins,
         mapping = aes(x = species,
                       y = log_mass,
                       color = island)) + 
  geom_boxplot()+ 
  geom_smooth(method = "lm")+
  labs(x = "Species", 
       y = "Log Body Mass (g)",
       color = "Island",
       title = "Log Body Mass of Female Penguin Species") +
  scale_colour_ghibli_d("MarnieMedium1") +
  theme_igray() +
  theme(axis.title = element_text(size = 20),
        panel.background = element_rect(fill = "linen"))
```





######################################################################

### Week 4b
### Data Wrangling: tidyr
### Created by: Haley Poppinga
### Created on: 2025_09_16
###########################################

NOTES

1. Intro to {tidyr}. 
* Separate and unite columns
* Pivot data between long and wide formats
2. Export csv with summary statistics. 


#### Review 
1. What function do you use to subset rows by some criterion?
* filter()

2. How do I add a new column to my data frame?
* mutate()


## Intro to the {tidyr} package (part of the TidyVerse)

New dataset!
We are going to use a subset of data from Silbiger et al. 2020 Proceedings of the Royal Society: B.

In the data folder for this week in Fall-2025, you will see two .csv files:

chemicaldata_manualua.csv
chem_data_dictionary.csv
Move both of them to the data folder of YOUR repo.

Explore the data dictionary to understand what you will be working with.

Basics: I collected biogeochemistry data at sites in Hawai'i with submarine groundwater discharge (fresh groundwater that flows into coastal oceans through cracks in the reef plate). We have data on nutrient and carbonate chemistry from two sites, two seasons, high and low tide, and during the day and night.



### Created by: Haley Poppinga
### Created on: 2025_09_16
###########################################

### Load libraries ########## 
```{r}
library(tidyverse)
library(here)
library(ggplot2)
```

### Read in data ###
```{r}
ChemData<-read_csv(here("Week_04","data", "chemicaldata_maunalua.csv"))
View(ChemData)
glimpse(ChemData)
```

There are several ways we can remove all the NAs in a data set
### Data Analysis ###
```{r}
#Remove NAs
ChemData_clean<-ChemData %>%
  filter(complete.cases(.)) #filters out everything that is not a complete row
View(ChemData_clean)
```

Separate function
We notice a silly column named Tide_time, which has tide (low/high) paired with time of day (day/night).  
Nyssa did NOT follow directions for tidy data from the first week of class... Now we have to fix it.
Fit is so just a column for time and just a column for tide

Notice that I have two bits of information: Low_Day is low tide during the day. It would be way easier to plot and analyze the data if we separate those into separate columns.  

hit ?separate to see all the things you can do, but the simplest use is:

separate(  
data = [data frame you are using],  
col = [column that you want to separate],  
into = [name of the new columns],  (of different things)
sep = [what are you separating by?]). 

Let's continue to build off our data frame

```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row, no NAs in data
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_" ) # separate by _ (an underscore, can be seperated by whatever is between them in data set e.g ., etc.)
head(ChemData_clean)
```

Notice this deletes the original column. If we wanted to keep it we would add remove = FALSE.
```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% 
  separate(col = Tide_time, 
           into = c("Tide","Time"), 
           sep = "_", 
           remove = FALSE) # keep the original tide_time column
head(ChemData_clean)
```


### Unite
What if we have the opposite problem and what to unite two columns into one?

Let's say we want a new column that unites site with zone column. B/c we want to make a plot that separates site and zone as the factor.  
We use the unite() function which has the same arguments as separate() and brings 2 columns together.  


What if we have the opposite problem and what to unite two columns into one?

Let's say we want a new column that unites site with zone. We use the unite() function which has the same arguments as separate()

```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>% # keep the original tide_time column
  unite(col = "Site_Zone", # the name of the NEW col
         c(Site,Zone), # the columns to unite
        sep = ".", # lets put a . in the middle
        remove = FALSE) # keep the original
head(ChemData_clean)
```
quotes for "Site_Zone" b/c telling it to make a column called that

### Pivoting the dataset between wide and long. 
Review. 

Wide data
* One observation per row and all the different variables are columns. 

Long data
* One unique measurement per row and all the info about that measurement in the same row. 

#### Pivoting between long and wide in R
* Wide to long: pivot_longer()
* Long to wide: pivot_wider()

Let's pivot our data so that it is in long format.
* We want one column with all the names of the biogeochemical parameters (i.e., NN, P, Si, etc)
* Paired with one column with all the values associated with those variables
* We want all the metadata (lat, long, tide, etc.) to be preserved in the correct order

Why is long format helpful?
* Easier to summarize using group_by(). You would only need to group by the variable name to get all summary statistics for every variable
* Easier to facet_wrap() by each variable if all the plots are going to be identical instead of making 10 individual plots

Nyssa's data is wide


### Pivot_longer()
Start with our ChemData_clean dataframe (note, you can also just pipe what we are about to do to your existing lines of code as well)

```{r}
ChemData_long <-ChemData_clean %>%
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. cols = columns you want to pull from This says select the temp to percent SGD cols,just biogeochem data)
               names_to = "Variables", # the names of the new cols with all the column names, rename it
               values_to = "Values") # names of the new column with all the values, extract just to values
View(ChemData_long)
```


What can we do with the long data set?
Let's calculate the mean and variance for all variables at each site

```{r}
ChemData_long %>%
  group_by(Variables, Site) %>% # group by everything we want, (variables and site which are the column names)
  summarise(Param_means = mean(Values, na.rm = TRUE), # get mean of all parameters, exclude NAs
            Param_vars = var(Values, na.rm = TRUE)) # get variance
## summarise() has grouped output by 'Variables'. You can override using the .groups argument.
# made easier to calculate mean/variance for all parameters in just 2 lines of code
```


Calculate mean, variance, and standard deviation for all variables by site, zone, and tide

Here is your starter code:
```{r}
ChemData_long %>%
  group_by(Variables, Site) %>% # group by everything we want 
  summarise(Param_means = mean(Values, na.rm = TRUE), # get mean 
            Param_vars = var(Values, na.rm = TRUE)) # get variance
```

```{r}
ChemData_long %>%
  group_by(Variables, Site, Zone, Tide) %>% # group by site, zone, and tide 
  summarise(Param_means = mean(Values, na.rm = TRUE), # get mean 
            Param_vars = var(Values, na.rm = TRUE), # get variance
            Param_sd = sd(Values, na.rm = TRUE)) # get standard deviation
```


Example using facet_wrap with long data
Create boxplots of every parameter by site
```{r}
ChemData_long %>%
  ggplot(aes(x = Site, y = Values))+ # raw values on y with site on x
  geom_boxplot()+
  facet_wrap(~Variables) #add in this to do facet as a function of variable
#makes boxplot for every single variable
# but all of the plots look the same because the scales don't work for every variable
```

```{r}
ChemData_long %>%
  ggplot(aes(x = Site, y = Values))+ 
  geom_boxplot()+ 
  facet_wrap(~Variables, scales = "free") #add in scales for each variable (min and max)
#free scales the x and the y 
# if only want to free x-axis --> free_x (keeps y same and changes x)
```


Let's say we got data in long format and need to convert it to wide. 
We use pivot_wider()
instead of names_to we use names_from
```{r}
ChemData_wide<-ChemData_long %>%
  pivot_wider(names_from = Variables, # column with the names for the new columns, Variables already exists in dataframe
              values_from = Values) # column with the values
View(ChemData_wide)
```



Let's calculate some summary statistics and export the csv file
Start from the beginning and work through our entire flow again, ending with data export.

Remove all NAs.
```{r}
ChemData_clean<-ChemData %>%
  drop_na()  #filters out everything that is not a complete row
View(ChemData_clean) #view with every single line added in console
```


Separate Tide_time into two columns.
```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE)
```
View(ChemData_clean)


Pivot the data longer
```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>%
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols
               names_to = "Variables", # the names of the new cols with all the column names
               values_to = "Values") # names of the new column with all the values
```
View(ChemData_clean)


Group by Variable, Site, Time and calculate means.
```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>%
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols  
               names_to = "Variables", # the names of the new cols with all the column names 
               values_to = "Values") %>% # names of the new column with all the values 
  group_by(Variables, Site, Time) %>%
  summarise(mean_vals = mean(Values, na.rm = TRUE))
```
View(ChemData_clean)



Convert back to wide
```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>%
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols  
               names_to = "Variables", # the names of the new cols with all the column names 
               values_to = "Values") %>% # names of the new column with all the values 
  group_by(Variables, Site, Time) %>% # wamt averages for day and night for each of sites for all variables, take pivot longer set and pipe to group by function and pipe to summarise --> columns for each of these variables with their mean 
  summarise(mean_vals = mean(Values, na.rm = TRUE)) %>%
  pivot_wider(names_from = Variables, # pivot so all variables are on the top
              values_from = mean_vals) # notice it is now mean_vals as the col name
```
View(ChemData_clean)



Export the csv file of your summary statistics using write_csv().
```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>%
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols  
               names_to = "Variables", # the names of the new cols with all the column names 
               values_to = "Values") %>% # names of the new column with all the values 
  group_by(Variables, Site, Time) %>% 
  summarise(mean_vals = mean(Values, na.rm = TRUE)) %>%
  pivot_wider(names_from = Variables, 
              values_from = mean_vals) # notice it is now mean_vals as the col name
  #write_csv(here("Week_04","output","summary.csv"))  # export as a csv to the right folder
```
## summarise() has grouped output by 'Variables', 'Site'. You can override using
## the .groups argument.



# Homework

Using the chemistry data:
- Create a new clean script
- Remove all the NAs
- Separate the Tide_time column into appropriate columns for analysis
- Filter out a subset of data (your choice)
- use either pivot_longer or pivot_wider at least once
- Calculate some summary statistics (can be anything) and export the csv file into the output folder
- Make any kind of plot (it cannot be a boxplot) and export it into the output folder
- Make sure you comment your code and your data, outputs, and script are in the appropriate folders

```{r}
ChemData_clean<-ChemData %>%
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>%
  pivot_longer(cols = Temp_in:percent_sgd, # the cols you want to pivot. This says select the temp to percent SGD cols  
               names_to = "Variables", # the names of the new cols with all the column names 
               values_to = "Values") %>% # names of the new column with all the values 
  group_by(Variables, Site, Time) %>% 
  summarise(mean_vals = mean(Values, na.rm = TRUE)) %>%
  pivot_wider(names_from = Variables, 
              values_from = mean_vals) # 
```









