---
title: "Week 06 Quarto Homework"
author: "Haley Poppinga"
date: today
format: 
  html:
    toc: true
    theme: sandstone
knitr: 
  opts_chunk:
    fig.path: "../Output/"
---
    
## Week 06 Homework

**Homework**  

* Take any of the data sheets you have already worked with in class and create at least one table and one figure in an Quarto file.  
* Create the most beautiful Quarto output you can in either an html, github doc, or pdf format. RENDER THE FILE
* Make sure you have headings with clear explanations of what you are doing.
* Practice using markdown formatting with bold, italics, and lists (use each of these at least once).
* Make sure your outputs and scripts are saved in the appropriate folders. Keep proper coding etiquette (figures need to be pub quality as usual and you need to comment your code).


### Load Libraries
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ghibli) 
library(beepr)
library(kableExtra)
library(tidyplots)
```


### Read Data

```{r}
#| message: false
#| warning: false

ChemData<-read_csv(here("Week_04","data", "chemicaldata_maunalua.csv"))
glimpse(ChemData)
```


### Make a Plot

This is looking at the relationship between penguin **pH**, **%SGD**, and **Season**.  
I wanted to try a density plot from ggplot to make the data visualization awesome. I'm still trying to understand this data in the right context, but it was good practice to struggle with the making of this plot. 

```{r}
#| warning: false
#| out-width: "70%"
#| fig-align: "center"
#| label: fig-chemdata
#| fig-cap: "This is a figure showing water chemistry data in Maunalua Bay by season."
#| echo: false
#| message: false
#| include: true
#| eval: true
#| dev: "png"
#| fig-format: png
#| fig-width: 7
#| fig-height: 3.5

ChemData_clean<-ChemData %>% #Nyssa's groundwater data
  drop_na() %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time column
           into = c("Tide","Time"), # separate it into two columns tide and time
           sep = "_", # separate by 
           remove = FALSE) %>% # keep the original tide_time column
  mutate(Season = factor(Season, levels = c("SPRING","FALL")),
         Tide = factor(Tide, levels = c("Low","High"))) 

# Stuff dealing with density plots
n_levels <- 6 # Number of filled contour bands
breaks <- c(0.25, 0.40, 0.55, 0.70, 0.85, 0.95, 0.995)  # above near-zero density
# original plot showed too much white space so added this to get a tighter view, set limits on axes
xlims <- quantile(ChemData_clean$pH, c(0.02, 0.98), na.rm = TRUE)
ytop  <- quantile(ChemData_clean$percent_sgd, 0.90, na.rm = TRUE)  


ChemData_clean %>%
 ggplot(aes(x = pH, y = percent_sgd)) + #density plot
  geom_density_2d_filled(contour_var = "ndensity", breaks = breaks) + 
  geom_density_2d(color = "white", linewidth = 0.3, breaks = breaks) +   # thin white contour lines on top (had to look up breaks stuff)
  facet_wrap(~ Season, nrow = 1) +
  labs(x = "pH",
       y = "%SGD",
       title = "pH and Submarine Groundwater Discharge (SGD) Density Regions by Season",
       fill = "Density Level") + # legend label
  coord_cartesian(xlim = xlims, ylim = c(0, 2.5)) + # zoom in and less white space
  scale_fill_brewer(palette = "Blues") + # colors to show pH better
  theme_minimal(base_size = 11) + # labels/theme for a pub visualization
  theme(panel.grid = element_blank(),
        plot.title = element_text(face = "bold",
                                  size = 12))
```
When you look @fig-chemdata we can see the density differences between pH and %SGD between Fall and Spring.



### Make a Table  

```{r}
#| warning: false
#| label: tbl-chemdata
#| tbl-cap: "Water Chemistry Data in Maunalua Bay"

ChemData_clean %>%
  drop_na() %>%
  pivot_longer(cols = c(pH, percent_sgd), #make data longer
               names_to = "Variable",
               values_to = "Value") %>%
  mutate(Variable = dplyr::recode(Variable, percent_sgd = "%SGD")) %>%  # rename variable
  group_by(Season, Variable) %>%
  summarise(Variable_mean = mean(Value, na.rm = TRUE)) %>% # get mean 
  kbl(col.names = c("Season","Variable","Mean")) %>% # make it a kable table and rename headers
  kable_classic_2() %>% # add a theme
  row_spec(1, bold = TRUE, color = "white", background = "blue") %>% # highlight row 3
  row_spec(3, bold = TRUE, color = "white", background = "orange") %>%
  kable_styling(full_width = FALSE) # make it less wide
```

@tbl-chemdata has information on the mean values for pH and %SGD by different seasons.


_Beep when the code is done running..._
```{r}
beep(10)
```


_Thank you Nyssa and Callie!_  








#### Extra  

_Some Extra Stuff_:  
To-Do before running code:  
.rs.restartR()  
rm(list = ls())  


Need to go to Office Hours with help rendering as a gfm format. It looked terrible when I did it.  
format: gfm  
prefer-html: true  

Also couldn't figure out how to change the background color of themes even using the Quarto website.  








