---
title: "Week_10 Lectures"
author: "Haley Poppinga"
format: 
  html: 
    toc: true
    theme: sandstone
---

## Week 10a Lecture: Functional Programming  

[Lecture Slides](https://raw.githack.com/OCN-682-UH/Poppinga/main/Week_10/Lectures/17_Functional-Programming.html#1)  
[Required Reading R4DS Ch. 25](https://r4ds.hadley.nz/functions.html)  
[Required Reading R4DS Ch. 8](https://r4ds.hadley.nz/workflow-help.html)  


### Introduction  

**Outline of class**  

Functional programming  

* Quiz!  
* When to write a function  
* How to make a custom function  

Homework: Practice writing functions.  


**Why functions?**  

* Anytime you have a task that you want to use more than once   
* Keeps your script clean  
* Avoids mistakes with copy and paste  
* Minimizes places where you need to update your code (copy and pasting code over and over makes it hard to change small things)    
* only have to change things once in a function

Anytime you have to do a task more then once...  
Let's use an example where we want to rescale data in multiple columns (value - min/(max - min))  


First set-up your script and create a dataframe of random numbers  


### Load Libraries  

```{r}
library(tidyverse)
library(here)
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
library(palmerpenguins)
library(PNWColors) # for the PNW color palette 
```


### Practice
```{r}
df <- tibble(
  a = rnorm(10), # draws 10 random values from a normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)
```
rnorm = randomly distributed numbers  
Basien modeling: use setseed() allows it to be consistent. Run before random number generator.  

Anytime you have to do a task more then once...  
Let's use an example where we want to rescale data in multiple columns (value - min/(max - min))  

*Rescale every column individually*  
```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)))
```
Replaces a withing the function to do whatever you tell it to.  



Now copy and paste the same code for each column without making any mistakes. 
```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))
```
Did I make a mistake?  
Yes-> after max()  

Correct:  
```{r}
df<-df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(b, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))
```


We can write a function for this  
```{r}
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
df %>%
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))
```



Let's break it down
There are three key steps to creating a new function:

1. You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.

2. You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).

3. You place the code you have developed in body of the function, a { block that immediately follows function(...).  
  * function() then {}

4. return tells us what values we want returned when you run the function  



```{r}
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE)) # value can be whatever name you want, 
  return(value) 
}
```
One argument here.  
rescale01 is now the function that you've created (e.g. read_csv)
A package is multiple functions put together to use in a library.  

Note: It is way easier to create a function when you already wrote the code rather than creating the function from scratch.


```{r}
mutate_all(rescale01)
```





Let's make a function to covert degrees fahrenheit to celcius.
What is the calculation for F to C?

Start out by writing the code you want to use:  
```{r}
temp_C<- (temp_F - 32)*5 / 9
```

Then figure out what are the different arguments and the inputs and outputs. 
In this case, input is farenheit and output is celsius.  

#### Let's make it a function!  

**Step 1: Name the function**  
```{r}
fahrenheit_to_celsius <- function() {
}
```


**Step 2: Put in the equation**  
```{r}
fahrenheit_to_celsius <- function() { 
  temp_C <- (temp_F - 32) * 5 / 9
}
```

**Step 3: Decide what the arguments are**  
```{r}
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9 
}
```
Note: want to make function names applicable to everyone.  

**Step 4: Decide what is being returned**  
```{r}
fahrenheit_to_celsius <- function(temp_F) { 
  temp_C <- (temp_F - 32) * 5 / 9 
  return(temp_C)
}
```
Note: put your functions at top of your script so it is available throughout the whole script.

**Step 5: Test it**  
```{r}
fahrenheit_to_celsius(32)
```

```{r}
fahrenheit_to_celsius(212)
```
Note: test with numbers that you know easily off the top of your head.  
  



##### Think, Pair, Share
Write a function that converts celcius to kelvin. (Remember Kelvin is celcius + 273.15).  

Code I want to use:  
temp_K<- (temp_C + 273.15)

```{r}

celsius_to_kelvin<-function(temp_C) # decide what the arguments are (step 3)
  {  # name the function (step 1)
  temp_K<- (temp_C + 273.15) # put in the equation (step 2)
  return(temp_K) # decide what is being returned (step 4)
}

celsius_to_kelvin(100)
```
Example use: calculate NCP or NCC


#### Making plots into a function  
Let's say you have a specifically formatted plot that you like and that you plan to use over and over again. By turning it into a function, you only have to code it once.  

Let's use our favorite penguin dataset and make a nice plot line plot.  

```{r}
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()
```
Note: can make a function like this to use the same theme throughout multiple plots.  

**Making plots into a function**  

Name and set-up the function.  
```{r}
myplot<-function(){ # copy and pasted code above into the body of the function here: 
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```


What do you think our arguments should be?  

* this funciton can change data, x, and y depending on what data you decide to use  
* 


Make the names broad so it can be applicable to several values
```{r}
myplot<-function(data, x, y){
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data = data, aes(x = x, y =y , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```


Test it:  

Well, shoot, I got an error.... why do we think that is?  
Why an error? 
The function does not know that those are within the data set. So we need to be explicit about what columns are.  
```{r}
myplot(data = penguins, x = body_mass_g, y = bill_length_mm)
```

Even though body_mass_g exists within the penguin dataframe, there is still no individual parameter called body_mass_g in our environment and R is confused.  

**There is a solution though from within the {rlang} package (a part of the tidyverse)!**  

* {rlang} uses what is literally called a "curly-curly" {{}} to help us assign variables that are column names in dataframes.  
* Make it easier to manipulate them within functions  


Let's add curly-curlies to the column names  
```{r}
myplot<-function(data, x, y){ 
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```
Use in a function any time you want to use a column within a tibble.  
{{}} is saying that you can put anything in x and anything in y  

Test again
```{r}
myplot(data = penguins, x = body_mass_g, y = bill_length_mm)
```

Test with new variables.  
Whatever continuous x or y and flip it out to whatever. 
Will always be the same colors to islands and same geom_smooth, etc.  
```{r}
myplot(data = penguins, x = body_mass_g, y = flipper_length_mm)
```




#### Adding defualts  

Let's say you want to create a default for the function to always default to the penguins dataset. You can set those directly in the function.  

```{r}
myplot<-function(data = penguins, x, y){ # going to always assume that data is penguins
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```


Defaults  
Now, you can just write...  
Dont need to write data because it is assuming the data is penguins.  
```{r}
myplot(x = body_mass_g, y = flipper_length_mm)
```


#### Layering the plot  
You can also layer onto your plot using '+' just like it is a regular ggplot to change things like labels.  
```{r}
myplot(x = body_mass_g, y = flipper_length_mm)+
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")
```

Note: can now manipulate plot in any way we already learned with the + signs  




**Add an if-else statement for more flexibility.**  
An aside on if-else statements....  

If this argument is true, then add something.  

Imagine you want a variable to be equal to a certain value if a condition is met. This is a typical problem that requires the if ... else ... construct. For instance:  
```{r}
a <- 4
b <- 5
```

Suppose that if a > b then f should be = to 20, else f should be equal to 10. Using if/else we:  
```{r}
if (a > b) { # my question
  f <- 20 # if it is true give me answer 1
    } else { # else give me answer 2
  f <- 10
}
```
If a > b make f assigned to 20, if it is anything else make f assigned to 10  

When I type f I get...  
```{r}
f
```
Because a is less than b in this case.  


**Back to plotting**  
Using if...else... we can make our function even more flexible.  
Let's say we want the option of adding the geom_smooth lines or not.  
We can create a variable that if set to TRUE add the geom_smooth, otherwise print without.  

First add a new argument for lines and make the default TRUE for ease.  
```{r}
myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```

If-else
Next, add an if-else statement
```{r}
myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
if(lines==TRUE){
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
} # curly for if statement
else{
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+ # no geom_smooth here
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
} # curly for else statement
} # curly for entire function
```



Test it
With lines
```{r}
myplot(x = body_mass_g, y = flipper_length_mm)
```

Without lines
```{r}
myplot(x = body_mass_g, y = flipper_length_mm, lines = FALSE)
```




Today's totally awesome R package  
{emokid}: Make your code more emo... you've used the praise package, but if you need to express your frustrations now you can let the world know you are sad

devtools::install_github("itsrainingdata/emokid")
For when you are really sad

library(emokid)
iamsad()
more nights of hugging my pillow, oh replaying memories. 😿
For when you are less sad

iamlesssad()
i can't dream anymore since you left. i miss you singing me to sleep. 😔




Homework
Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using Quarto with at least 2 examples per function showing its flexibility. Due, as always, at 1pm on Tuesday.

_________________________________________________________________________________


## Week 10b Lecture: Getting Help in R    

[Lecture Slides](https://raw.githack.com/OCN-682-UH/Fall_2025/main/Week_10/21_GettingHelp.html#1)  
[Required Reading R4DS Ch. 8](https://r4ds.hadley.nz/workflow-help.html)  


### Introduction  

**Outline of class**  

Functional programming  

* Quiz!  
* When to write a function  
* How to make a custom function  
































