---
title: "Week 10 Homework"
author: "Haley Poppinga"
format: 
  html: 
    toc: true
    theme: sandstone
---

### Introduction  

**Homework**  
Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using Quarto with at least 2 examples per function showing its flexibility. Due, as always, at 1pm on Tuesday.  


**Three key steps to creating a new function:**  

1. You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.    

2. You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).   

3. You place the code you have developed in body of the function, a { block that immediately follows function(...).  
  * function() then {}  

4. return tells us what values we want returned when you run the function   



### Load Libraries  
```{r}
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(ghibli)
library(forcats)
library(stringr)
library(ggmap)
library(sf)
```


### Read Data  
```{r}
#| warning: false
#| message: false
chemdata<-read_csv(here("Week_10","Data","chemicaldata_maunalua.csv")) # Maunalua chem data
```

```{r}
#| warning: false
#| message: false
glimpse(chemdata)
```

### Clean the Data  
```{r}
#| warning: false
#| message: false

chemdata_clean<-chemdata %>%
  drop_na() %>% #filters out everything that is not a complete row
# Pivot Longer
  pivot_longer(cols = c(pH, Salinity, NN, Silicate, Phosphate, TA, percent_sgd), # pH, salinity, and nutrients
               names_to = "Variable",    # columns for those
               values_to = "Value") %>%  # measurement and all metadata
# Summarise Means and Variances
  group_by(Site, Lat, Long, Variable) %>% # group by zone and variables
  summarise(parameter_mean = mean(Value, na.rm = TRUE), # get mean 
            parameter_variance = var(Value, na.rm = TRUE)) # get variance
#View(chemdata_clean)
```




## Make a Function 1 (Not a Plot)  

Start out by writing the code you want to use:  

variation<-(variance / ((mean)^2)

Then figure out what are the different arguments and the inputs and outputs. 
For example, I want to use Lubarsky et al. 2018 calculation of **pH variance**.  
```{r}
#| warning: false
#| message: false

variation_equation<-function(variance, mean) { # name the function and add arguments
  variation<-variance / ((mean)^2) # put in equation for pH variation
  return(variation) # output the result
}
```


### Function 1 Example 1  
Calculate pH variation of experimental locations in Wailupe from chemdata:  
```{r}
#| warning: false
#| message: false

pH_wailupe_variation<-chemdata %>% 
  drop_na(pH) %>% # drop NAs for pH
  group_by(Site, Lat, Long) %>% 
  filter(Site == "W") %>% 
  summarise(pH_mean = mean(pH, na.rm = TRUE),
            pH_variance = var(pH, na.rm = TRUE)) %>% 
  mutate(variation = variation_equation(pH_variance, pH_mean))

pH_wailupe_variation
```


### Function 1 Example 2  
Calculate all parameter variation in both sites from chemdata:  
```{r}
#| warning: false
#| message: false

chemdata_variation<-chemdata_clean %>% 
  mutate(variation = variation_equation(parameter_variance, parameter_mean)) # calculates variation for all parameters
head(chemdata_variation)
```




________________________________________________________________________________


## Make a Function 2 (Plot)  
Make a map plot of latitude and longitude coordinates of experimental locations, like that in Lubarsky et al. 2018 or Silbiger et al. 2020:  

*Note: You will need your own Google Maps API key sorry*  
```{r}
#| warning: false
#| message: false

# register_google(key = "") # use your own API in between the "" in R console

# Site maps for any latitude/longitude data

chemdata_map<-function(data, x, y,
                       maptype = "satellite", #Google map type
                       zoom = 17, # zoom into center
                       pt_size = 2,
                       pt_col = "yellow") 
{  # name function and determine arguments

# palette<-scale_color_ghibli_d("MarnieMedium2", direction = -1) # nice colors if I am creating a legend for NEP or pH perhaps


# Pull longitude and latitude automatically from any column data as numeric vectors
lon<-pull(data, {{ x }}) # uses dplyr
lat<-pull(data, {{ y }})

# center map at mean coordinates
map_center<-c(lon = mean(lon, na.rm = TRUE),
              lat = mean(lat, na.rm = TRUE))

basemap<-get_googlemap(center = map_center, 
                 maptype = maptype, zoom = zoom) #base map from Google Maps

ggmap(basemap) +
  geom_point(data = data, mapping = aes(x = {{x}}, y = {{y}}), # uses rlang to signify it is within dataset
                       size = pt_size, color = pt_col) + # constants outside aes()
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        panel.grid = element_blank())
}
```



### Function 2 Example 1   
Map of experimental locations for Wailupe:  
```{r}
#| warning: false
#| message: false

chemdata %>% 
  filter(Site =="W") %>% 
  chemdata_map(Long, Lat) + 
  labs(x = "Longitude", y = "Latitude", title = "Wailupe Experimental Locations") # could add this into the function when I have more time to figure that out
```


### Function 2 Example 2  
Map of experimental locations for Black Point:  
```{r}
#| warning: false
#| message: false

chemdata %>% 
  filter(Site =="BP") %>% 
  chemdata_map(Long, Lat, zoom = 18) + 
  labs(x = "Longitude", y = "Latitude", title = "Black Point Experimental Locations")
```






Extra:  
Future things to do when more time: make a scale bar.  

See the way [Nyssa](https://github.com/njsilbiger/MaunaluaSEM/blob/159fcb935bf467edb70d9cfb035e7200b76ab0e9/mapplot_noAPI.R) did it.  

```{r}
# ggsave(here("Week_10","Output","functionplot1_HW10.png"))
# Empty your environment before you start working
# Restart R
# .rs.restartR()
# remove whole list in enviroment
# rm(list = ls())
```








