---
title: "Goodplot/Badplot"
author: "Haley Poppinga"
date: today
format: 
  html:
    toc: true
    theme: sandstone
---

## Introduction  

#### Assignment:  

**Instructions**  
Good plot Bad plot lab – competition (there is a winner and the winner gets a prize)  

Due October 21, 2025  

Homework:  
Your homework is to create two versions of similar plots, using any dataset you choose. You must create the plots using ggplot2, post-processing with Adobe Illustrator or similar is not allowed. This homework is a contest, and the winner will win a prize to be disclosed.  

Please submit your homework by posting a Quarto document containing the plots, the code you used to create them, and your discussion of the plot within the Quarto doc to your GitHub folder. Create an independent folder for your goodplot-badplot. Follow all best practices for file structure that we already learned in class. Your document MUST be rendered.  

 
The first plot should be as bad as possible. Based on your reading from [Claus Wilke’s](https://clauswilke.com/dataviz/) and [Healy's textbook](https://socviz.co/index.html#preface) describe the many ways that a plot can become misleading, difficult to interpret, or just plain ugly, and you should use all of the tools that you can. Please try to make me cry, a person who cares a great deal about high-quality data visualization.  

You should explain in text all of the ways in which the bad plot is bad. Bullet points are fine but clearly articulate each principle of data visualization that you have broken.  

The second plot should be based on the same or similar data to the first plot, but it should be good. Write briefly about why this is a successful plot.  

 

_Badplot rules:_  

* All plots must be made in ggplot2, with code visible in the Quarto document.  
* The data must be visible.  
* The good plot and the bad plot should be based on the same, or similar, data.  
* I will post the results on Social Media. Please notify me if you would prefer not to have your plots posted publicly. That is fine, but I need to know. The community waits for this every year and they love it.  
 

This assignment is 20 points:  
5 points - follow best practices for code/file organization   
5 points  - creativity for good plot  
5 points - creativity for bad plot  
5 points - explanation of why plots either do (good plot) or don't (bad plot) follow best practices for data visualization  

[Inspiration for a great place to get data for this assignment](https://github.com/rfordatascience/tidytuesday)

Check: is it bad practice to do bad plot and good plot on same file?



### Load Libraries
```{r}
library(tidyverse)
library(here)
library(ggplot2)
library(tidytuesdayR)
library(ghibli) 
library(beepr)
library(lubridate)
library(dplyr)
```

### Read Data  
```{r}
#| message: false
#| warning: false

scoobydoo <- tt_load("2021-07-13") # loads Tidy Tuesday Scooby Doo data
head(scoobydoo)
```

```{r}

```


## Bad Plot  


```{r}
piechart<-ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) # basic pie chart
  theme_void() + # remove background, grid, numeric labels
  geom_text(aes(label = group), position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette="Set1")
  
p2 <- ggplot(df, aes(x, y)) + # second plot
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  theme_bw()

p3 <- ggplot(df, aes(x, y)) + # third plot
    geom_point(alpha = 0.3) +
    geom_smooth(se = FALSE) +
    scale_x_continuous(limits = c(0, 0.5)) +
    scale_y_continuous(limits = c(-0.3, 0.6)) +
    theme_bw()

p2 + annotation_custom(ggplotGrob(p2), xmin = 1, xmax = 3, 
                       ymin = -0.3, ymax = 0.6) # second plot with third plot on top
```




























