---
title: "My awesome quarto"
author: "Nyssa Silbiger"
date: today
format: 
  html:
    toc: true
    theme: superhero 
---

# Introduction

Today we are going to work with the penguin data to learn *Quarto*.  

We are also learning basic Quarto.  Here are all the ways I can change an [HTML format Quarto Doc](https://quarto.org/docs/reference/formats/html.html)

# Load Libraries
First, we will load the important libraries.

```{r}
#| message: false # this turns off the messages

library(here)
library(tidyverse)
library(palmerpenguins)

```

# Analysis

Next, we need to look at the dataframe  
```{r}
glimpse(penguins) # look at the Penguin data
```

# Plot
We will make a plot looking at bill length versus bill depth

```{r}
#| warning: false

penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color = species))+
  geom_point()+
  geom_smooth(method = lm)+
  labs(x = " Bill Length (mm)",
       y = "Bill Depth (mm)")+
  theme_bw()
```

Here are some built in [themes](https://quarto.org/docs/output-formats/html-themes.html) for fun