---
title: "Week 07 Homework"
author: "Haley Poppinga"
date: "October 7, 2025"
format: 
  html:
    toc: true
    theme: sandstone
---

## Week 07 Homework

**Homework**  
You need to make a map using Quarto with one of the following datasets from past Tidy Tuesdays. Make sure, as usual, that you scripts and output are saved appropriately.  



### Load Libraries
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(here)
library(ggplot2)
library(ghibli) 
library(beepr)
library(lubridate)
library(dplyr)
library(maps)
library(mapdata)
library(mapproj)
library(ggspatial)
library(ggmap)
```



### Read Data
```{r}
#| message: false
#| warning: false

meteorites<-read_csv(here("Week_07","Data","meteorites.csv")) # meteorites data
glimpse(meteorites)
```

_Data for the entire World_  
```{r}
#| message: false
#| warning: false

# get data for the entire world
world<-map_data("world")  # map of the world
head(world) # not necessary but just to see

```

This meteorite data comes from Meteoritical Society by way of NASA. It includes where and when meteroties fell.  

Here is the link to the [Meteroites](https://github.com/rfordatascience/tidytuesday/tree/main/data/2019/2019-06-11) dataset! Tidy Tuesdays.   


### Make a Map  

**Clean the Data**  
```{r}
#| message: false
#| warning: false
meteorites_clean<- meteorites %>%
  drop_na() %>%
  mutate(fall = factor(fall, levels = c("Found","Fell")),
         long = if_else(long > 180, long - 360, long))
```

**Style and Fix the Map**  
```{r}
#| message: false
#| warning: false
bbox <- c(left = -179.99, bottom = -80, right = 179.99, top = 80)   # cap north at ~80°
worldmap <- get_map(location = bbox, maptype = "satellite", crop = FALSE)
# Get base layer
```

_Plot the Map reference_  
```{r}
#| message: false
#| warning: false
#| out-width: "70%"
#| fig-align: "center"
#| fig-height: 4
#| fig-width: 4

# plot it
ggmap(worldmap)
```

**Meteorite Map**  
```{r}
#| warning: false
#| out-width: "70%"
#| fig-align: "center"
#| label: fig-meteorites
#| fig-cap: "This is a figure showing where meteorites fell or were found around the world."
#| echo: false
#| message: false
#| include: true
#| eval: true
#| dev: "png"
#| fig-format: png

ggmap(worldmap) +
  geom_point(data = meteorites_clean,
             aes(x = long, y = lat, color = fall), # map by where meteorites fell and where found
                 size = 0.25) + # size of points
  scale_color_ghibli_d("MarnieMedium2", direction = -1) + # nice colors
  annotation_scale(bar_cols = c("orange","grey"),
                   location = "br",
                   pad_x = unit(0.25, "cm")) + # put the bar on the bottom left and make the colors yellow and grey
  annotation_north_arrow(location = "tl", 
                         style = north_arrow_fancy_orienteering,
                         height = unit(1, "cm"), width = unit(1, "cm"),
                         pad_y = unit(0.25, "cm")) + # add style and make north arrow smaller
  coord_sf(crs = 4326) +  # scale bar to work, needs to be in this typical coordinate reference system for a GPS (WGS84)
  labs(x = "Longitude",
       y = "Latitude",
       title = "Meteorites Fell or Found Around the World",
       subtitle = "Data from the Meteoritical Society by way of NASA",
       color = "Fell or Found") + # legend label
  theme_minimal(base_size = 11) + # labels/theme for a pub visualization
  theme(panel.grid = element_blank(),
        plot.title = element_text(face = "bold", size = 12))

# ggsave(here("Week_07","Output","meteoritesplot_HW7.png"))
```
@fig-meteorites shows the locations of where meteorites fell (green) and where they were found found (yellow).



Note:  
Spent too much time trying to figure out how to include the Pacific. Gave up. 
