---
#| title: "How to Ask for Help Online for R (and Use LLMs Wisely)"
#| author: "Dr. Nyssa Silbiger"
#| date: "`r format(Sys.Date())`"
#| lang: en
#| format:
#|   html:
#|     toc: true
#|     toc-title: "On this page"
#|     toc-depth: 3
#|     theme: flatly
#|     code-fold: true
#|     code-summary: "Show code"
#|     code-tools: true
#|     code-line-numbers: true
#|     number-sections: true
#|     df-print: paged
#|     link-external-newwindow: true
#| execute:
#|   echo: true
#|   warning: false
#|   message: false
#| filters: [rmarkdown]
#| editor: visual
---

<!--
Optionally render as slides by switching to revealjs:
format:
  revealjs:
    theme: [simple, default]
    slide-number: true
    toc: true
    toc-depth: 2
    incremental: true
    chalkboard: true
    code-line-numbers: true
-->


## 🎯 Learning Objectives

By the end of this session, you will be able to:

- Write clear, reproducible help requests for R (forums, GitHub, Slack, AI tools).
- Create minimal **reprex** examples and include the right diagnostics.
- Apply the **Do's and Don'ts** of using ChatGPT (and other LLMs) for R.
- Compare ChatGPT, Gemini, Claude, and Copilot for common R workflows.

::: callout-important
### Expectation for this course (OCN-682)
When you post for help (class Slack/Posit Community/Stack Overflow), include: a **title**, a **reprex**, **sessionInfo()**, **expected vs. actual output**, and **what you tried**. Submissions without these will be returned for revision.
:::


## 🧩 Setup

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(fig.retina = 2, dpi = 96)
```

```{r}
#| label: packages
#| message: false
needed <- c("tidyverse", "reprex")
new <- setdiff(needed, rownames(installed.packages()))
if (length(new)) install.packages(new)
lapply(needed, require, character.only = TRUE)
```


## 🧠 Why Asking for Help is a Core Research Skill

> Good questions get good answers.

You will often seek help via **Posit Community**, **Stack Overflow**, **GitHub Issues**, **Reddit r/rstats**, and **LLMs** (ChatGPT, Gemini, Claude) as well as lab Slack. Clear communication accelerates your research and helps others help you.


## 📐 The Anatomy of a Good Help Request

### ✅ Include These

1. **Concise problem statement**
2. **Minimal reproducible example (reprex)**
3. **Session information**
4. **Expected vs. actual output**
5. **What you tried**

### Reprex template

```{r}
#| label: reprex-template
#| eval: false
# minimal data
library(tidyverse)
df <- tibble(year = c(2020, 2021), richness = c(5, 8))

# problematic code
p <- ggplot(df, aes(year, richness)) + geom_line()
print(p)

# your R session
sessionInfo()

#### This is the example chatGPT produced and the code actually runs so it is not "problematic". One example on how ChatGPT is not always correct.
```

Run `reprex::reprex()` to produce a clean, copy-paste-able post for forums.


### Session info (paste at end of post)

```{r}
#| label: session-info
#| eval: false
sessionInfo()
```


## 🚫 Common Mistakes (and Fixes)

| ❌ Don’t | ✅ Do |
|---|---|
| "My code doesn’t work. Help!!!" | Provide exact **error text** + **minimal code** |
| Screenshots of code | Paste code as **text** (fenced code blocks) |
| Vague titles | Specific titles (e.g., *Error with `mutate()` after join*) |
| Missing data context | Include a **small sample** (`dput(head(df))`) |
| Blame the tool | Focus on **debugging steps** and intent |


## 💬 Where to Ask for Help

| Platform | Best For | Tone/Style |
|---|---|---|
| **Posit Community** | Beginner–intermediate R | Friendly, educational |
| **Stack Overflow** | Precise code Qs | Formal, concise, minimal |
| **GitHub Issues** | Package bugs & feature requests | Structured, technical |
| **Reddit r/rstats** | Casual discussion | Conversational |
| **LLMs (ChatGPT, Gemini, Claude, Copilot)** | Brainstorming, quick feedback | Conversational, iterative |


## 🤖 Do's & Don'ts of Using ChatGPT for R

:::: {.panel-tabset}
### ✅ Do

- Use it as a **thinking partner**, not an oracle.
- Provide **full context** (goal, reprex, error).
- **Iterate**: ask for refinements (e.g., “dplyr-only version”).
- Ask for **explanations** (concepts + why the fix works).
- **Validate** code locally; read docs.
- Keep **privacy**: redact sensitive/unpublished data.

### 🚫 Don't

- Don’t paste confidential data or credentials.
- Don’t copy code blindly; avoid cargo-culting.
- Don’t ask **vague** questions without a reprex.
- Don’t skip learning—use AI as a tutor.
- Don’t rely on hallucinated packages/functions—**verify**.
::::


## ⚖️ LLM Comparison for R Workflows

> **Note:** Capabilities evolve; verify current features in your environment.

| Feature | ChatGPT (GPT‑5 Thinking) | Gemini 1.5 Pro | Claude 3.5 Sonnet | GitHub Copilot |
|---|---|---|---|---|
| **Strengths** | Strong reasoning; good R examples; multi-step debugging | Multimodal inputs; code comprehension | Clear explanations; safe style | Inline completions in RStudio/VS Code |
| **Weaknesses** | May hallucinate APIs; needs verification | R package edge-cases | May truncate long code | Limited broader context |
| **Best For** | Debugging, Quarto/Rmd authoring, doc writing | Explaining figures/images; brainstorming | Design choices; doc reviews | Speeding up repetitive code |
| **Privacy** | Organizational controls available | Workspace integrations | Team policy controls | Enterprise options via Microsoft |

::: callout-note
**Tip:** Try the **same prompt** across tools and compare: clarity, correctness, runnable code, and explanation depth.
:::


## 🧪  Activity: “Help Me Help You” 

1. **Bad post → Good post**: Convert a vague help request into a high‑quality one (use the rubric below).
2. **Prompt the LLMs**: Use identical prompts with ChatGPT, Gemini, and Claude.
3. **Compare**: Which result is most correct, runnable, and well‑explained?

### What to think about
- Title specific and informative 
- Minimal reprex provided 
- Error text included 
- Expected vs. actual clearly stated 
- Steps tried listed 


## 🧰 Prompt & Post Templates

:::: {.panel-tabset}
### Template: Forum Post (Posit/Stack Overflow)

```text
**Title:** `mutate()` fails after `left_join()` with duplicated keys

**Goal**: Summarize survey counts by site and year, then plot a trend.

**Problem**: Error `object 'n' not found` when running `mutate()` after `left_join()`.

**Reprex:**

```r
library(tidyverse)
set.seed(1)
raw <- tibble(site = rep(LETTERS[1:3], each = 4),
              year = rep(2020:2021, times = 6),
              species = sample(c("Acropora","Porites"), 12, TRUE))

sum_by <- raw |>
  count(site, year, name = "n")

lookup <- tibble(site = LETTERS[1:3], region = c("N","C","S"))

joined <- left_join(sum_by, lookup, by = "site")

joined |>
  mutate(prop = n / sum(n))
```

**Expected vs. actual**: expected proportions by site; actual error above.

**Session info**:

```r
sessionInfo()
```

**What I tried**: `group_by()` before `mutate()`, `ungroup()`, and `dplyr::summarise()`.
```

### Template: LLM Prompt (ChatGPT/Gemini/Claude)

```text
You are an R tutor. Diagnose and fix the error in my reprex. Explain **why** it occurs and give two alternative fixes (tidyverse‑native and base R). Provide runnable code only.

<PASTE REPREX HERE>
```
::::


## 🔍 Debugging Patterns to Try (before posting)

- **Minimal data**: `head(df, 20)`
- **Check types**: `str(df)`, `is.factor(x)`, `glimpse(df)`
- **Isolate**: run each pipe step line by line
- **Name conflicts**: prefer `dplyr::filter()` over `stats::filter()`
- **Fresh session**: restart R, `sessionInfo()`


## 🛡️ Ethics, Reuse, and Citations

- Redact sensitive or unpublished data before posting anywhere.
- Attribute ideas/code adapted from others (forums, blogs, AI).
- For manuscripts, follow journal policy on AI assistance and include a disclosure if required.


## 📚 Quick Links

- Posit Community: *How to make a great reprex*
- Stack Overflow: [R tag](https://stackoverflow.com/questions/tagged/r)
- `reprex` docs: https://reprex.tidyverse.org/
- GitHub Issues templates (many packages include a bug report template)


## ✅ TChecklist

- [ ] Clear title
- [ ] Minimal reprex
- [ ] Error text
- [ ] Expected vs. actual
- [ ] `sessionInfo()`
- [ ] What you tried
- [ ] Privacy check


---

